---
title: "DataAnalysis"
author: "Sam Gene"
date: "January 23, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Part 2 of Week three assignment

####Read in measurments.csv that was created in part 1 of assignment for week 3
```{r}
MData<-read.csv("Data\\Measurments.csv")
head(MData)
str(MData)
```

####Change columns names to shorter 3-letter abbreviations
```{r}
library(dplyr)
MData<-MData %>% rename(Spc=Species,LbW=Limb.Width,UnW=unitsW,LbL=Limb.Length,UnL=unitsL)
head(MData)
```

####Add a Volume column
#####But first convert the units so that they are all the same
```{r}
MData<-MData %>% mutate(LbW = ifelse(UnW=="mm",LbW*0.1,LbW),LbL = ifelse(UnL=="mm",LbL*0.1,LbL),UnW = replace(UnW, UnW=="mm", "cm"),UnL = replace(UnL, UnL=="mm", "cm"))
MData

```

#####Add volume column
```{r}
MData<-MData %>% mutate(Vol=(pi*((LbW/2)^2))*LbL)
head(MData)
```

####Add columns for groups

```{r}
MData
Class<-c(rep(c("Bivalve"),4),rep(c("Amphibian"),6))
MData<-cbind(MData,Class);MData
```
####Summarize mean and sd of each group
```{r}
SumDat<-MData %>% group_by(Class) %>% summarize(MeanVol=mean(Vol),SDVol=sd(Vol),MeanWidth=mean(LbW),SDWidth=sd(LbW),MeanLength=mean(LbL),SDLength=sd(LbL))
SumDat
```
####Plot Width and length
```{r}
library(ggplot2)
qplot(LbW, LbL, data = MData,xlab="Width (cm)",ylab="Length (cm)", 
      main= "Mussel and amphibian limb length and width",alpha=I(0.5), shape=I(18),colour=Class, size=I(5))+
    scale_colour_manual(values=c("seagreen","steelblue"))+
    labs(colour= "Class")+
  theme_classic()
```

####Plot only those that are above average volume

```{r}
summarize(MData,MeanVol=mean(Vol))
```
```{r}
AboveAv<-MData %>% filter(Vol>1846.532)
qplot(LbW, LbL, data = AboveAv,xlab="Width (cm)",ylab="Length (cm)", 
      main= "Above average mussel and amphibian limb length and width",alpha=I(0.5), shape=I(19),colour=Class, size=I(5))+
    scale_colour_manual(values=c("forestgreen","navy"))+
    labs(colour= "Class")+
  theme_classic()
```

# Add a able of the above average volume individuals

Species | Class | Limb Width (cm) | Limb Length (cm) | Volume (cm^3)
-----------|--------|-----|-----|-----
Epioblasma triquetra | Bivalve | 5.597 |	91.13 | 2242
Pseudacris triseriata | Amphibian |	7.388	| 127.0	|	5446
Anaxyrus americanus	| Amphibian | 10.53	|	91.02	|	7932
Lithobates pipiens | Amphibian |	5.465 | 110.3	|	2588
